{\bf Systematic Dynamic Race Condition Detection}

### 2013-04-15 landslide-lecture/lecture.tex
Revise slide deck for 2013

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2012, April 6}

@@

\begin{frame}[fragile]{Case Study}
	\begin{center}
	\begin{verbatim}
	int thread_fork()
	{
	    thread_t *child = spawn_new_thread();
	    add_to_runqueue(child);
	    return child->tid;
	}
	\end{verbatim}
	\end{center}

@@

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{slaughter cho2} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

@@

\begin{frame}{Execution Tree Exploration}
	Important point: When does a branch end?
	\begin{itemize}
		\item All threads run to completion, or
		\item A bug is detected % Say: more on this later.
	\end{itemize}
	{\bf Backtracking}:
	\begin{itemize}
		\item Identify a decision point to choose differently
		\item Reset machine state and start over
		\item Replay test from the beginning, with different choices
	\end{itemize}
\end{frame}

@@

	\linegap
	One extreme: Decide at every instruction
	\begin{itemize}
		\item Good news: Will find every possible race condition.
		\item Bad news: Runtime of test will be impossibly large.
	\end{itemize}
	\linegap

	Other extreme: Nothing is a decision point
	\begin{itemize}
		\item Good news: Test will finish quickly.
		\item Bad news: Only one execution was checked for bugginess.
		\item Bad news: No alternative interleavings explored.
		\begin{itemize}
			\item Makes ``no race found'' a weak claim.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{About The Project}
	5th year MS since June 2011 % Say: planning to finish this may, hopefully

@@

	\linegap
	Checkpointing/backtracking via Simics bookmarks
\end{frame}

@@

\begin{frame}{Identifying Bugs}
	Landslide can {\em definitely discover}:
	\begin{itemize}
		\item Kernel panics % Say: note: the better your asserts are..!
		\item Deadlock
		\item Use-after-free / double-free
	\end{itemize}
	Landslide can {\em reasonably suspect}:
	\begin{itemize}
		\item Memory leak
		\item Progress sense (halting problem)
	\end{itemize}
\end{frame}

@@

	Mutual benefit
	\begin{itemize}
		\item Landslide may help you find bugs
		\item You may help Ben evaluate his thesis project
	\end{itemize}
\end{frame}

@@

\begin{frame}{Annotating Your Kernel}
	Step 1
	\begin{center}
		\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	Your kernel needs to say when certain events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap

	\framebox{Time estimate: 40 minutes}
\end{frame}

@@

	Fill in two implementation-dependent C functions in Landslide ($\le$10 lines)
	% Say: Which depend on your kernel's specific implementation
	\linegap

	\framebox{Time estimate: 60 minutes}
\end{frame}

\begin{frame}{Configuring Decision Points}
	Landslide automatically identifies a minimal set of decision points.
	\begin{itemize}
		\item It might find bugs.
		\item It might overlook more fine-grained interleavings.
	\end{itemize}
	With help from you, it could find more.
	\begin{itemize}
		\item Optional annotation: \texttt{tell\_landslide\_decide()}
		\item Hints to where a context switch should be forced.
		\item Inside every call to \texttt{mutex\_lock}\ldots
	\end{itemize}
\end{frame}

@@

\begin{frame}{In Which Ben Offers Help - Warning}
	\textbf{If you are already struggling, this will not ``save'' you.}
	\begin{itemize}
		\item False-negatives: not guaranteed to find races at all
		\item Research-quality: possibly difficult to integrate with your kernel
		\item Finishing the kernel project is more important.
	\end{itemize}
	You should:
	\begin{itemize}
		\item Be expecting an A or B
		\item Be able to turn in without late days (if you really had to)
		\item Be looking for\ldots
			\begin{itemize}
				\item That ``one pesky race''
				\item A race that stress-testing missed
				\item Or just familiarity with a new technique
			\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{In Which Ben Offers Help}
	User study next week, starting Monday

	\linegap
	Expect to spend:
	\begin{itemize}
		\item Up to 4 hours, just to try it out.
		\item 6-8 hours, if you find a bug and track it down.
		\item More, for multiple bugs or the truly dedicated\ldots
	\end{itemize}
	\linegap

>>>>

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2013, April 14}

@@

\begin{frame}[fragile]{Case Study}
	\begin{center}
	\begin{verbatim}
	    int thread_fork()
	    {
	        thread_t *child = construct_new_thread();
	        add_to_runqueue(child);
	        return child->tid;
	    }
	\end{verbatim}
	\end{center}

@@

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{slaughter cho2} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	\linegap
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

@@

\begin{frame}{Execution Tree Exploration}
	Important point: When does a branch end?
	\begin{itemize}
		\item All threads run to completion, or
		\item A bug is detected % Say: more on this later.
	\end{itemize}
	\linegap
	{\bf Backtracking}:
	\begin{itemize}
		\item Identify a decision point to choose differently
		\item Reset machine state and start over
		\item Replay test from the beginning, with different ``decisions''
	\end{itemize}
\end{frame}

@@

	\linegap
	One extreme: Decide at every instruction
	\begin{itemize}
		\item {\em Good news}: Will find every possible race condition.
		\item {\em Bad news}: Runtime of test will be impossibly large.
	\end{itemize}
	\linegap

	Other extreme: Nothing is a decision point
	\begin{itemize}
		\item {\em Good news}: Test will finish quickly.
		\item {\em Bad news}: Only one execution was checked for bugginess.
		\item {\em Bad news}: No alternative interleavings explored.
		\begin{itemize}
			\item Makes ``no race found'' a weak claim.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{About The Project}
	5th year MS since June 2011 % Say: planning to finish this may, hopefully
	My MS thesis project (June 2011-2012)

	\linegap
	Now I'm a Ph.D. student\dots

@@

	\linegap
	Checkpointing/backtracking via Simics bookmarks (reverse execution)
\end{frame}

@@

\begin{frame}{Identifying Bugs}
	Landslide can {\em definitely discover}:
	\begin{itemize}
		\item Kernel panics % Say: note: the better your asserts are..!
		\item Deadlock
		\item Use-after-free / double-free
	\end{itemize}
	\linegap
	Landslide can {\em reasonably suspect}:
	\begin{itemize}
		\item Memory leak
		\item Infinite loop (halting problem)
	\end{itemize}
\end{frame}

@@

	Mutual benefit
	\begin{itemize}
		\item Landslide may help you find bugs
		\item You help Ben evaluate his research
	\end{itemize}
\end{frame}

@@

\begin{frame}{Annotating Your Kernel}
	Step 1
	\begin{center}
		\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	Your kernel needs to say when certain events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
\end{frame}

@@

	Fill in two implementation-dependent C functions in Landslide ($\le$10 lines)
	% Say: Which depend on your kernel's specific implementation
\end{frame}

\begin{frame}{Configuring Decision Points}
	Landslide automatically identifies a minimal set of decision points.
	\begin{itemize}
		\item It might find bugs.
		\item It might overlook more fine-grained interleavings.
	\end{itemize}
	\linegap
	With help from you, it could find more.
	\begin{itemize}
		\item Optional annotation: \texttt{tell\_landslide\_decide()}
		\item Hints to where a context switch should be forced.
		\item Inside every call to \texttt{mutex\_lock}\ldots
	\end{itemize}
\end{frame}

\begin{frame}{And Hopefully\dots}
	\begin{center}
		\includegraphics[width=\textwidth]{found_a_bug.png}
	\end{center}
\end{frame}

@@

\begin{frame}{Last Year's Results}
	\begin{itemize}
		\item Worked with 5 groups
		\begin{itemize}
			\item \dots{}of which 4 put in enough time to make Landslide work
		\end{itemize}
		\linegap
		\item {\bf Investment}: about 2 to 3 hours
		\begin{itemize}
			\item Average 112 minutes doing required instrumentation
			\item Average 35 minutes configuring extra decision points \& interpreting ``found a bug'' output
		\end{itemize}
		\pause
		\linegap
		\item {\bf Return}: All 4 groups found bugs in their kernels.
		\begin{itemize}
			\item All groups found deterministic bugs (e.g., use-after-free)
			\item 2 groups found race conditions (i.e., Landslide needed to backtrack)
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{In Which Ben Offers Help - Warning}
	\textbf{If you are already struggling, this will not ``save'' you.}
	\begin{itemize}
		\item False-negatives: not guaranteed to find races at all
		\item Research-quality: possibly difficult to integrate with your kernel
		\item Finishing the kernel project is more important.
	\end{itemize}
	\pause
	\linegap
	This is for you if:
	\begin{itemize}
		%\item Be expecting an A or a B
		\item You have already submitted
		\item You are using late days, but willing to work over Carnival
		%\item Be able to turn in without late days (if you really had to)
		\item You are taking p3extra, but you already pass the hurdle
		\item You are looking for\ldots
			\begin{itemize}
				\item That ``one pesky race''
				\item A race that stress-testing missed
				\item Or just familiarity with a new technique
			\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{In Which Ben Offers Help}
	User study this week, by appointment (\texttt{bblum@cs.cmu.edu})

	\linegap
	Expect to spend:
	\begin{itemize}
		\item Up to 3 hours, just to try it out.
		\item 4-6 hours, if you find a bug and track it down.
		\item More, for multiple bugs or the truly dedicated\ldots
	\end{itemize}
	\linegap
